Lensing in the Schwarzschild Spacetime Applicable to Neutron Stars

### 2022-09-01 src/tex/ms.tex
Fix doppler sign error, and add more description of e/e_surf terms.

For a photon striking the surface of the neutron star with momentum $p^\mu\left(
e, l \right)$, the energy measured by the observer fixed to the surface is 
\begin{equation}
    -e_\mathrm{surf} = g_{\mu\nu} \frac{\dd \chi^\mu}{\dd \lambda} p^\nu;
\end{equation}
equivalently, for a photon \emph{emitted} from the surface with energy
$e_\mathrm{surf}$ in the local rest frame (e.g.\ thermally from particles on the
surface), the above equation relates the emitted energy to the energy as
measured by observers at infinity ($e$).

Plugging in Eqs.\ \eqref{eq:pt-final}, \eqref{eq:pr-final},
\eqref{eq:ptheta-final}, and \eqref{eq:pphi-final}, we obtain the relation
between the energy observed at infinite distance ($e$) and the emitted energy at
the surface ($e_\mathrm{surf}$) for a ray with impact factor $b = x \bNS$
hitting the surface of the neutron star at colatitude $\theta_0$:
\begin{multline}
    \frac{e}{e_\mathrm{surf}} = 1 + y \vRot \sin \theta_0 + \left( y^2 - \frac{1}{2} \right) \vRot^2 \sin^2 \theta_0 \\ 
    - \frac{1}{\rNS} \left( 1 - \left( y^2 - \frac{1}{2} \right) \vRot^2 \sin^2 \theta_0 \right) + \ldots, 
\end{multline}
where $y = x \sin \theta_0$ is the ``equatorial'' component of the
(dimensionless) impact factor.

>>>>

For a photon striking the surface of the neutron star with momentum $p^\mu\left(
e, l \right)$, the energy measured by the observer fixed to the surface is 
\begin{equation}
    -e_\mathrm{surf} = g_{\mu\nu} \frac{\dd \chi^\mu}{\dd \lambda} p^\nu.
\end{equation}
If we reverse the spatial components of the momentum, then for a photon
\emph{emitted} from the surface with energy $e_\mathrm{surf}$ in the frame fixed
to the stellar surface (e.g.\ thermally from particles on the surface), the
above equation relates the emitted energy to the energy as measured by observers
at infinity ($e$).

Plugging in Eqs.\ \eqref{eq:pt-final}, \eqref{eq:pr-final},
\eqref{eq:ptheta-final}, and \eqref{eq:pphi-final}, remembering to reverse the
spatial components of the momentum, we obtain the relation between the energy
observed at infinite distance ($e$) and the emitted energy at the surface
($e_\mathrm{surf}$) for a ray with impact factor $b = x \bNS$ hitting the
surface of the neutron star at colatitude $\theta_0$:
\begin{multline}
    \label{eq:energy-relation}
    \frac{e}{e_\mathrm{surf}} = 1 - y \vRot \sin \theta_0 + \left( y^2 - \frac{1}{2} \right) \vRot^2 \sin^2 \theta_0 \\ 
    - \frac{1}{\rNS} \left( 1 - \left( y^2 - \frac{1}{2} \right) \vRot^2 \sin^2 \theta_0 \right) + \ldots, 
\end{multline}
where $y = x \sin \theta_0$ is the ``equatorial'' component of the
(dimensionless) impact factor.  (Recall $x, y < 0$ for points west of the
meridian and $x, y > 0$ for points to the east; $\sin \theta_0 > 0$ always
because $0 < \theta_0 < \pi$.)

The velocity-independent term in Eq.\ \eqref{eq:energy-relation} gives the
gravitational redshift; the term linear in velocity gives the doppler shift from
the rotational motion (blueshift---increased energy---west of the meridian,
redshift---decreased energy---for points to the east); and the term quadratic in
the velocity comes from relativistic beaming.
